### 2013-07-30 Tex/chapters/results.tex
ausführlichere Überleitung
Korrekturen vorheriger Irrtümer

\subsection{Short Analysis of the Autocorrelation Time  $\tau$}
\label{ssec:results:autocorr}
    % Sollte das hier rein?
    Before the results are presented, a short analysis of the
    autocorrelation time \(\tau\) is given, to illustrate the benefits
    of the used Wolff cluster algorithm.\\
    In \ref{fig:autocorr}\subref{fig:autocorr:temperatures}
    the autocorrelation time for the magnetization per site \(m\) at \(\sigma=0\) is plotted.
    \begin{figure}[htbp]
        \centering
        \subfigure[in Dependence on the Temperature $T$][]
        {
            \label{fig:autocorr:temperatures}
            \includegraphics[width=0.45\textwidth]{plots/autocorrT}
        }
        \subfigure[in Dependence on the System Size $L$][]
        {
            \label{fig:autocorr:powerLaw}
            \includegraphics[width=0.45\textwidth]{plots/autocorrL}
        }
        \caption[The Autocorrelation Time $\tau$]
        {
            Dependece of the autocorrelation time $\tau$ on
            \subref{fig:autocorr:temperatures} the temperature $T$ for
                $\sigma=0$ and
            \subref{fig:autocorr:powerLaw} the system size $L$ (errorbars
                are the standard error determined by bootstrapping)
        }
        \label{fig:autocorr}
    \end{figure}
    The plateau at low temperatures is easy to understand considering the
    effects of the Wolff algorithm. At low temperatures it flips nearly every
    spin in every step, thus the correlation drops to zero after one sweep.
    Alternatively the parallel tempering algorithm swaps spin configurations
    with possibly different signs, thus having the same effect.
    Also note that the maximum of \(\tau\) is not at \(T_c\) but at a higher
    temperature. The cause is probably the effectiveness of the Wolff cluster
    algorithm at \(T_c\), hence \(\tau \approx 1\) at \(T_{c}\).
    Obviously the autocorrelation time \(\tau\) increases with the system
    size. In fact it obeys a power law \(\tau \propto \L^z\), which is
    the expected behavior of a dynamical exponent \(z\) as mentioned in
    section \ref{sssec:eqtime}.
    In fig. \ref{fig:autocorr}\subref{fig:autocorr:powerLaw} \(\tau\) is plotted
    over \(L\). Note that the exponent \(z'=0.64(2)\) determined by this plot is
    not comparable to the known critical exponent \(z=0.25\) \cite{NewmanBarkema1999}
    for the Wolff cluster algorithm. On the one hand the number of sweeps
    in which \(\tau\) is measured in this thesis are more than a standard
    Metropolis sweep to which \(z\) normally corresponds, thus \(z'\) would
    need to be normalized by some factor \(>1\), because each sweep
    alongside the \(N\) Metropolis flips also a cluster of \(\ge 1\) spin is
    flipped. On the other hand the \(\tau\) used for the fit are not the
    \(\tau\) at \(T_{c}\) but the maximum \(\tau\) of all \(T\).
    Nevertheless it is interesting, that it obeys a power law
    \(\tau \propto L^{z'}\) just the same. This suggests, that \(\xi \gtrsim L\)
    is satisfied at the temperature of maximum \(\tau\) and that \(z'\) is a
    valid upper bound of \(z\).
    This proofs the effectiveness of the Wolff cluster algorithm at
    criticality.\\
    This is of course only a rough analysis of existing data which was
    generated for another purpose. For a more in depth inspection, one
    would perform the simulation at the critical point, which will be
    determined in the next sections.

\subsection{Finite Size Effects}
\label{ssec:finitesize}
    The aim of this thesis is to find the critical temperature \(T_c\)
    of the disordered Ising model in dependence of the disorder parameter
    \(\sigma\). At \(T_c\) the mean absolute magnetization \(\avg{|m|}\) of
    the system will show a steep decline to zero and the susceptibility
    \(\chi = \frac{N}{T}\brac{\avg{m^2}-\avg{m}^2}\)
    will diverge. In fig. \ref{fig:smeared_out}\subref{sfig:smeared_out:meanM_L}
    it is easy to see that the steep decline of \(\avg{|m|}\)
    occurs at lower temperatures \(T\) for higher
    disorder parameters \(\sigma\) (here with a RNG).
    \begin{figure}[htbp]
        \centering
        \subfigure[Dependence of the Phase Transition on $\sigma$][]
        {
            \label{sfig:smeared_out:meanM_L}
            \includegraphics[width=0.45\textwidth]{plots/Mean_M_L_128}
        }
        \subfigure[Example for Finite Size Effects][]
        {
            \label{sfig:smeared_out:meanM}
            \includegraphics[width=0.45\textwidth]{plots/meanM}
        }
        \caption[Phase Transition and Finite Size Effects]
        {
            \subref{sfig:smeared_out:meanM_L}: Effects of the disorder
            parameter $\sigma$ on the phase transition
            for an underlying RNG with $L=128$.\\
            \subref{sfig:smeared_out:meanM}: Effects of different system
            sizes at \(\sigma = 0\). Dotted lines are guides to the eye.
        }
        \label{fig:smeared_out}
    \end{figure}\\
    But in the shown plots there occurs no steep decline to zero, but a
    smooth decline. \(\avg{\abs{m}}(T_c) = 0\) is only present in infinite
    systems, hence no computer simulation will show the exact behavior in the
    thermodynamic limit. It will always show some \emph{finte size effects}.
    These finite size effects cause a "smearing out" of the phase
    transition. This is stronger for smaller system sizes, as is visible
    in fig. \ref{fig:smeared_out}\subref{sfig:smeared_out:meanM}\footnote{See the appendix \ref{appendix:finiteSizeEffects} for a similar figure for the specific heat.}.
    Clearly the \(L=16\) curve is much less steep than the \(L=128\) curve.\\
    Despite of this one can obtain \(T_c\) by finite size scaling
    methods \cite[S. ??]{NewmanBarkema1999} which also yield the critical
    exponents.

>>>>

\subsection{Short Analysis of the Autocorrelation Time  $\tau$}
\label{ssec:results:autocorr}
    % Sollte das hier rein?
    Before the results are presented, a short analysis of the
    autocorrelation time \(\tau\) is given, to illustrate the benefits
    of the used Wolff cluster algorithm.\\
    In \ref{fig:autocorr}\subref{fig:autocorr:temperatures}
    the autocorrelation time for the magnetization per site \(m\) at \(\sigma=0\) is plotted.
    \begin{figure}[htbp]
        \centering
        \subfigure[in Dependence on the Temperature $T$][]
        {
            \label{fig:autocorr:temperatures}
            \includegraphics[width=0.45\textwidth]{plots/autocorrT}
        }
        \subfigure[in Dependence on the System Size $L$][]
        {
            \label{fig:autocorr:powerLaw}
            \includegraphics[width=0.45\textwidth]{plots/autocorrL}
        }
        \caption[The Autocorrelation Time $\tau$]
        {
            Dependece of the autocorrelation time $\tau$ on
            \subref{fig:autocorr:temperatures} the temperature $T$ for
                $\sigma=0$ and
            \subref{fig:autocorr:powerLaw} the system size $L$ (errorbars
                are the standard error determined by bootstrapping)
        }
        \label{fig:autocorr}
    \end{figure}
    The plateau at low temperatures is easy to understand considering the
    effects of the Wolff algorithm. At low temperatures it flips nearly every
    spin in every step, thus the correlation drops to zero after one sweep.
    Alternatively the parallel tempering algorithm swaps spin configurations
    with possibly different signs, thus having the same effect.
    Also note that the maximum of \(\tau\) is not at \(T_c\) but at a higher
    temperature. The cause is probably the effectiveness of the Wolff cluster
    algorithm at \(T_c\), hence \(\tau \approx 1\) at \(T_{c}\).
    Obviously the autocorrelation time \(\tau\) increases with the system
    size. In fact it obeys a power law \(\tau \propto \L^z\), which is
    the expected behavior of a dynamical exponent \(z\) as mentioned in
    section \ref{sssec:eqtime}.
    In fig. \ref{fig:autocorr}\subref{fig:autocorr:powerLaw} \(\tau\) is plotted
    over \(L\). Note that the exponent \(z'=0.64(2)\) determined by this plot is
    not comparable to the known critical exponent \(z=0.25\) \cite{NewmanBarkema1999}
    for the Wolff cluster algorithm. On the one hand the number of sweeps
    in which \(\tau\) is measured in this thesis are more than a standard
    Metropolis sweep to which \(z\) normally corresponds, because each sweep
    alongside the \(N\) Metropolis flips also a cluster of \(\ge 1\) spin is
    flipped. On the other hand the \(\tau\) used for the fit are not the
    \(\tau\) at \(T_{c}\) but the maximum \(\tau\) of all \(T\).
    Nevertheless it is interesting, that it obeys a power law
    \(\tau \propto L^{z'}\) just the same. This suggests, that \(\xi \gtrsim L\)
    is satisfied at the temperature of maximum \(\tau\), which is plausible,
    because the peak in fig. \ref{fig:autocorr}\subref{fig:autocorr:powerLaw}
    is near \(T_{c}\).\\
    Anyhow, fig. \ref{fig:autocorr}\subref{fig:autocorr:powerLaw} alone
    proofs the effectiveness of the Wolff cluster algorithm at criticality.\\
    This is of course only a rough analysis of existing data which was
    generated for another purpose. For a more in depth inspection, one
    would perform the simulation at the critical point, which will be
    determined in the next sections.

\subsection{Finite Size Effects}
\label{ssec:finitesize}
    The aim of this thesis is to find the critical temperature \(T_c\)
    of the disordered Ising model in dependence of the disorder parameter
    \(\sigma\). At \(T_c\) the mean absolute magnetization \(\avg{|m|}\) of
    the system will show a steep decline to zero and the susceptibility
    \(\chi = \frac{N}{T}\brac{\avg{m^2}-\avg{m}^2}\)
    will diverge. In fig. \ref{fig:smeared_out}\subref{sfig:smeared_out:meanM_L}
    it is easy to see that the steep decline of \(\avg{|m|}\)
    occurs at lower temperatures \(T\) for higher
    disorder parameters \(\sigma\) (here with a RNG).
    \begin{figure}[htbp]
        \centering
        \subfigure[Dependence of the Phase Transition on $\sigma$][]
        {
            \label{sfig:smeared_out:meanM_L}
            \includegraphics[width=0.45\textwidth]{plots/Mean_M_L_128}
        }
        \subfigure[Example for Finite Size Effects][]
        {
            \label{sfig:smeared_out:meanM}
            \includegraphics[width=0.45\textwidth]{plots/meanM}
        }
        \caption[Phase Transition and Finite Size Effects]
        {
            \subref{sfig:smeared_out:meanM_L}: Effects of the disorder
            parameter $\sigma$ on the phase transition
            for an underlying RNG with $L=128$.\\
            \subref{sfig:smeared_out:meanM}: Effects of different system
            sizes at \(\sigma = 0\). Dotted lines are guides to the eye.
        }
        \label{fig:smeared_out}
    \end{figure}\\
    But in the shown plots there occurs no steep decline to zero, but a
    smooth decline. \(\avg{\abs{m}}(T_c) = 0\) is only present in infinite
    systems, hence no computer simulation will show the exact behavior in the
    thermodynamic limit. It will always show some \emph{finte size effects}.
    These finite size effects cause a "smearing out" of the phase
    transition. This is stronger for smaller system sizes, as is visible
    in fig. \ref{fig:smeared_out}\subref{sfig:smeared_out:meanM}\footnote{See the appendix \ref{appendix:finiteSizeEffects} for a similar figure for the specific heat.}.
    Clearly the \(L=16\) curve is much less steep than the \(L=128\) curve.\\
    Despite of this one can obtain \(T_c\) by finite size scaling
    methods \cite[S. ??]{NewmanBarkema1999} which also yield the critical
    exponents, which is performed in the next section. Also in \ref{ssec:binderIntersections}
    an easier approach, which yields better values for \(T_{c}\), is presented.
